---
title: "Voting in the United Nations General Assembly"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 the variables are:

 Name                 Description
 -------------------- ----------------------------------------------------------
 `CountryName`        The name of the country
 `idealpoint`         Its estimated ideal point
 `Year`               The year for which the ideal point is estimated
 `PctAgreeUS`         The percentage of votes that agree with the US on the same issue
 `PctAgreeRUSSIA`     The percentage of votes that agree with Russia/the Soviet Union on the same issue



how the distribution of state ideal points has changed since the end of communism:

1. Plot the distribution of ideal points separately for 1980 and 2000 - about ten years before
   and after the fall of the Berlin Wall, respectively. 
   -Add median to each plot as a vertical line.  
```{r question1}
newset <- subset(unvoting$idealpoint, unvoting$Year == 1980 & unvoting$Year <= 2000 )
newset2 <- subset(unvoting$idealpoint, unvoting$Year == 2000 )
hist(newset, main="the distribution of ideal points in 1980", xlab = "Ideal Point", border="blue", col="pink")
summary(newset)
med <- median(newset)
abline(v = med, col = "blue", lwd = 2)

hist(newset2, main="the distribution of ideal points in 2000", xlab = "Ideal Point", border="blue", col="pink")
summary(newset)
med <- median(newset)
abline(v = med, col = "blue", lwd = 2)
```

2. How do the two distributions differ? 

   - the two differ in that countries seem a bit more polarized in 2000 than in 1980. You can see this in the way that most countries seem to be centered around the median in 1980 whereas in 2000, there bars are less centralized. This implies increased polarization. 

3. Use the `quantile` function to quantify the patterns you identified.
  Given the the results, one can see that the percentiles are more spread out. The numbers are less orderly as compared to 1980 where we can see a bit more symmetry in the quantiles.

```{r question1}
print(1980)
quantile(newset, probs = seq(0, 1, 0.25), na.rm = FALSE, names = TRUE)
print(2000)
quantile(newset2, probs = seq(0, 1, 0.25), na.rm = FALSE, names = TRUE)
```


how the number of countries voting with the US has changed over time.

1. Plot the average percent agreement with US across all counties over time.  Also, add the average percent agreement with Russia as another line for comparison.  
```{r question2}

us_avg_percent <- tapply(unvoting$PctAgreeUS, unvoting$Year, mean)
russia_avg_percent <- tapply(unvoting$PctAgreeRUSSIA, unvoting$Year, mean)

## US agree
plot(names(us_avg_percent), us_avg_percent, col = "red", type = "l",
xlim = c(1946, 2017), ylim = c(0, 1), xlab = "Year", main = "Avg agreement with all countries over time", ylab = "Avg percent agreement", pch = 18)

## Russia agree
lines(names(russia_avg_percent), russia_avg_percent, col = "blue")

# Add a legend to the plot
legend("topleft", legend=c("US", "Russia"),
       col=c("red", "blue"), lty = 1:2, cex=0.8)
```
2. Does the US appear to be getting more or less isolated over time, as compared to Russia?
  The Us appears US seems to be getting more isolated over time as compared to Russia. I say this seeing the percent of countries that agree with the United States decrease while Russia stays a few percentage points higher consistantly after the 1970s. 


```{r question2}

us_big_fan <- tapply(unvoting$PctAgreeUS, unvoting$CountryName, mean)
russia_big_fan <- tapply(unvoting$PctAgreeRUSSIA, unvoting$CountryName, mean)

print("US")
tail(sort(us_big_fan),5)
print("Russia")
tail(sort(russia_big_fan),5)

```
3. What are some countries that are consistently pro-US? 
Countries that pro - US consistently include Israel, Taiwan, United Kingdom, Palau, and of course, the USA.

4. What are the most pro-Russian countries?  
The most pro-Russian countries include Ukraine, Belarus, Czechoslovakia, German Democratic Republic, and of course, Russia. 

5. Give a brief substantive interpretation of the results.
These are countries that both respectively have had the most history with and shared culture. In the case of Russia, many of those countries are former soviet states. Likewise for the USA, we see the United Kingdom from which the country was established and other places such as Taiwan and Israel that are dependent on USA politics for their legitimacy in world politics.


1. plot the evolution of the two countries' ideal points over time. Add the yearly median ideal
  point of all countries. 
```{r question3}

question3_subset_US <- subset(unvoting, unvoting$CountryName == "United States of America")
question3_subset_Russia <- subset(unvoting, unvoting$CountryName == "Russia" )


Q3_US <- tapply(question3_subset_US$idealpoint, question3_subset_US$Year, mean)
Q3_R <- tapply(question3_subset_Russia$idealpoint, question3_subset_Russia$Year, mean)

all_ideal <- tapply(unvoting$idealpoint, unvoting$Year, median)
summary(all_ideal)

# US agree
plot(names(Q3_US), Q3_US, col = "red", type = "l",
xlim = c(1946, 2017), ylim = c(-3, 4), main = "Evolution of ideal points over time", pch = 18, ylab="Ideal points", xlab = "Time (years)")
# Russia agree
lines(names(Q3_R), Q3_R, col = "blue")
#median
lines(names(all_ideal), all_ideal, col="green", pch=1, type = 'p')

# Add a legend to the plot
legend("topleft", legend=c("US", "Russia"),
       col=c("red", "blue"), lty = 1:2, cex=0.8)



```
How might the results of this analysis modify (or not) your interpretation of the previous analysis?
These results solidifies my analysis that other countries are more closely aligned and voting with Russia.  However, that does not necesarily suggest that countries are moving closer *because* of Russia, rather that Russia has moved closer to what the rest of the world believes. We can see this with the increasingly similar frequency of ideal points between Russia and the rest of the other countries. The rest of the world votes more similarily to Russia than they do the US, supporting the claim that the US has become more isolated.


 The former Soviet Union countries are: Estonia, Latvia, Lithuania, Belarus, Moldova, Ukraine, Armenia, Azerbaijian, Georgia, Kazakhstan, Kyrgyzstan, Tajikistan, Uzbekistan, and Russia.The `%in%` operator, which is used as `x %in% y`, may be useful.  This operator returns a logical vector whose element is`TRUE` (`FALSE`) if the corresponding element of vector
  `x` is equal to a value contained in vector `y`.  Focus on the most recently available UN data from 2012 and plot each Post-Soviet Union state's ideal point against the percentage of its votes that agree with the United States. 
  
  Compare the post Soviet Union states, within the same plot, against the other countries. 
```{r question4}

former_countries <- c("Estonia", "Latvia", "Lithuania", "Belarus", "Moldova", "Ukraine", "Armenia", "Azerbaijan", "Georgia", "Kazakhstan", "Kyrgyzstan", "Tajikistan", "Uzbekistan", "Russia")

unvoting$former_countries <-ifelse(unvoting$CountryName %in% former_countries, 1, 0)
un_2012 <- subset(unvoting, unvoting$Year == 2012)

plot(un_2012$idealpoint[un_2012$former_countries == FALSE], un_2012$PctAgreeUS[un_2012$former_countries == FALSE], type = 'p', main = "Former states and us agreement/ ideal point", pch = 1,col="pink", ylab = "", xlab = "")
lines(un_2012$idealpoint[un_2012$former_countries == TRUE], un_2012$PctAgreeUS[un_2012$former_countries == TRUE], type = 'p', pch = 1,col="green")

legend("topleft", legend=c("Other countries", "Former States"),
       col=c("pink", "green"), lty = 1:2, cex=0.8)

```

   observations:
 - Former Soviet countries are seem to be more centralized than other countries in 2012, leaning neither very far left or very far right when compared to non Soviet states. They seem to retain similar opinions concerning ideal points which might be due to their shared history. That being said, there are some liberal states hovering close to the '1' on the x axis that are worth noting. 
  

how the median ideal points of Soviet/post-Soviet countries and all other countries has varied
  over all the years in the data. 
  
1. Plot these median ideal points by year. indicate 1989, the year of the fall of the Berlin

```{r question5}

post_su <- subset(unvoting, former_countries == 1)
non_su <-  subset(unvoting, former_countries == 0)

y1 <-tapply(non_su$idealpoint, non_su$Year, median)
y2 <-tapply(post_su$idealpoint, post_su$Year, median)

plot(names(y1), y1, type = 'l', main = "median ideal points", pch = 16, col="pink", ylim = c(-3,3), xlab = "year", ylab = "ideal point")
lines(names(y2), y2, type = 'l', pch = 1,col="green")
abline(v = 1989, col = "blue")
legend("topleft", legend=c("former states", "other", "fall of SU"),
       col=c("pink", "green", "blue"), lty = 1:2, cex=0.8)

```
observations:

The former states have become less likely to use ideal points after the fall of the soviet union as compared to other countries that seemed to spike in their usage of ideal points after that time. This suggests a major difference politically between former soviet states and other countries in the world, especially in terms of un voting trends.
  

apply the `k`-means clustering algorithm to ideal points and the percentage of votes agreeing with the US. Initiate the algorithm with just two centroids and visualize the results separately for 1989 and 2012. 
```{r question6}

ideal_points89 <- cbind(unvoting$PctAgreeUS[unvoting$Year == 1989], unvoting$idealpoint[unvoting$Year == 1989])
ideal_points12 <- cbind(unvoting$PctAgreeUS[unvoting$Year == 2012], unvoting$idealpoint[unvoting$Year == 2012])

# k-means with 2 clusters
k89four.out <- kmeans(ideal_points89, centers = 2, nstart = 5)
k12four.out <- kmeans(ideal_points12, centers = 2, nstart = 5)

#  plotting the results using the labels and limits defined earlier
lab = "Ideal Points"
plot(ideal_points89, col = k89two.out$cluster + 1, xlim = c(-.5, 1.5), ylim = c(-2.75, 3), main = "1989", xlab = lab, ylab = lab)
#  plotting the centroids
points(k89four.out$centers, pch = 8, cex = 2)
# 2012
plot(ideal_points12, col = k12four.out$cluster + 1, xlim =  c(-.5, 1.5), ylim = c(-2.75, 3), main = "2012", xlab = lab, ylab = lab)
points(k12four.out$centers, pch = 8, cex = 2)


```
2. Observations: Following the end of communism, countries that were formerly part of the Soviet Union have become much more ideologically diverse. Is this also true of the world as a whole? do countries still divide into two ideological factions? 
  - It seems that it is true that the rest of the world has become more ideologically diverse. We can see this how many of the points are more spread out than we see in 1989. However, there do still seem to be two factions, those who vote 0 to negative ideal points and those that vote 0+ ideal points.
  
  

We saw earlier that the median post-Soviet country joined the liberal cluster after the fall of communism.

1. Identify the countries which were in the non-liberal cluster in 1989 but belonged to the liberal cluster in 2012. Again, the `%in%` operator may be useful.  Then, do the reverse, so that you can see which countries exited the liberal cluster.
```{r question1}

cluster2012 <- table(Country = unvoting$CountryName[unvoting$Year == 2012], cluster = k12four.out$cluster)

margin.table(cluster2012,2)
cluster2012[ "cluster"][]


print(cluster2012)

```

there are 131 in the first cluster and  62 countries in the second cluster. 

